---
title: "Penguins analysis"
author: "R.G. Thomas"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
geometry: "left=3cm,right=5cm,top=2cm,bottom=2cm"
output:
  pdf_document:
    toc: true
    number_sections: true
    highlight: tango
    keep_tex: true
    includes:
      in_header: "~/shr/preamble.tex"
---

```{r include=F, echo=F}
library(pacman)
p_load(palmerpenguins, tidyverse, knitr)

opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE, fig.width = 4,
  fig.height = 2.1, results = "asis", dev = "pdf"
)
```

# Introduction

We can work with the dataset `penguins` included in the package `palmerpenguins`
```{r }
library(palmerpenguins)
```

One naive approach is to split the dataset and do three separate
analyses:

```{r }
df1 <- split(penguins, penguins$species)
```

The R package `purrr` provides a straightforward method to conduct the
analyses with a single command. Assume the set of data tables are
contained in a list of dataframes.  Also assume the analysis is a simple
visualization of a potential linear association between two features,

```{r }
df1 <- split(penguins, penguins$species)

foo <- function(df, z) {
  df |> ggplot(aes(x = bill_length_mm, y = flipper_length_mm)) +
    geom_point(aes(color = island), alpha = .5) +
    geom_smooth() +
    scale_color_manual(values = c("purple", "green", "red")) +
    theme_bw() +
    labs(
      title = paste(z, " Penguin Anatomy Comparison"), x = "Flipper length",
      y = "Bill length", color = "Island"
    )
  plotfile_name <- paste0(z, ".pdf")
  ggsave(plotfile_name)
  cat(paste0("\\includegraphics[]{", plotfile_name, "}"), "\n")
  cat(paste0("\\captionof{figure}{Scatterplot for species: ", z, "}"), "\n")
  cat("\\vspace{1cm}", "\n")
}

bar <- df1 |> map2(names(df1), foo)
```
